Automatically Splitting a Two-Stage Lambda Calculus

### 2015-07-10 paper/relatedwork.tex
related work

@@

Algorithms for stage splitting have appeared in the literature under the names 
{\em pass separation}, {\em binding time separation}, {\em data specialization}, 
{\em program bifurcation}, and as various unnamed compiler passes.

In the partial evaluation community, stage splitting has been investigated 
mostly as a preprocessing step for self-applied partial evaluators. 
For instance,~\cite{Mogensen89a} and~\cite{DeNiel91} defined splitting algorithms
for programs consisting of a set of first-order recursive functions.
Those works use a grammar-based binding-time specification that's 
relatively complicated compared to the type system of \lang.
This complication mostly arises from the decision to start with an untyped language:
one must first find a way to say ``argument $x$ will be a tuple'' before one can say
``the first component of $x$ is available at the second stage,'' 
and so much of their binding-time specification just amounts to a type system.
Formulations more like \lang\ (namely the two-level system of~\cite{NN92-twolevel}) were known at the time,
but~\cite{Mogensen89a} found these unsatisfactory because they could not describe data structures with mixed stage.
Davies essentially solved this limitation in~\cite{davies96} (pending our further tweak with $\curr$),
and the question of stage splitting as we have presented it has been ripe ever since.
Binding time analysis for a static higher-order language was visited in \cite{Mogensen89b},
but the implications for splitting were not explicitly considered.

>>>>

\TODO move this

In the partial evaluation community, metastatic staging transforms (called \emph{program bifurcation} and \emph{binding time separation}) 
similar to splitting have been investigated 
as a preprocessing step necessary for simplifying the specialization phase of self-applied partial evaluators. 
For instance, Mogensen\,\cite{Mogensen89a} and Niel et al.\,\cite{DeNiel91} defined transformations for a set of first-order recursive functions.

Those works operate on the results of binding time analysis

and so much of their binding-time specification just amounts to a type system.
Formulations like \lang\ (namely the two-level system of~\cite{NN92-twolevel}) were known at the time,
but~\cite{Mogensen89a} found these unsatisfactory because they could not describe data structures with mixed stage.

Davies essentially solved this limitation in~\cite{davies96}

Binding time analysis for a static higher-order language was visited in \cite{Mogensen89b}.

@@

Algorithms for stage splitting have appeared in the literature under the names 
{\em pass separation}\,\cite{JS86-staging} and {\em data specialization}\,\cite{knoblock96}.
